Hydra Battles and Cie

### 2021-08-25 doc/part-hydras.tex
Alectryon: 75% of chapter6
Alectryon: 85% of chapter6

\begin{Coqsrc}
Lemma L_finite : forall i k :nat,  L_ i k = (i+k)%nat.  
(* Proof by induction on i, using L_zero_eqn and L_succ_eqn *)

Lemma L_omega : forall k, L_ omega%e0 k = S (2 * k)%nat.
(* Proof using L_finite and L_lim_eqn *)
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma L_ge_S alpha : alpha <> Zero -> S <<= L_ alpha.

Theorem L_correct alpha : L_spec (cnf alpha) (L_ alpha).
\end{Coqsrc}

@@

Theorem \texttt{battle\_length\_std } of
Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html\#battle_length_std}{Hydra.Hydra\_Theorems} relates the length of standard battles with the functions $L_\alpha$.

\begin{Coqsrc}
Theorem battle_length_std (alpha : E0)  :
  alpha <> Zero ->
  forall k, (1 <= k)%nat ->
            battle_length standard k (iota (cnf alpha))
                         (L_ alpha (S k) - k).
\end{Coqsrc}

@@

\begin{Coqsrc}
Equations H'_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  H'_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Limitb alpha) :=
          { | left _ =>  H'_ (Canon alpha (S i))  i ;
            | right notlimit =>  H'_ (Pred alpha) (S i)}}.

Solve All Obligations with auto with E0.
\end{Coqsrc}

\begin{Coqsrc}
Lemma H'_eq1 : forall i, H'_ Zero i = i.
Proof.
  intro i; now rewrite H'__equation_1. 
Qed.

Lemma H'_eq2 alpha i : Succb alpha ->
                       H'_ alpha i = H'_ (Pred alpha) (S i).

Lemma H'_eq3 alpha i : Limitb alpha ->
                      H'_ alpha i =  H'_ (Canon alpha (S i)) i.

Lemma H'_succ_eqn  alpha i :  
            H'_ (Succ alpha) i = H'_ alpha (S i).
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H'_Fin : forall i k : nat,  H'_  (Fin i) k = (i+k)%nat.
Proof with eauto with E0.
  induction i.
  - intros; simpl Fin; simpl; autorewrite with H'_rw E0_rw ... 
  - intros ;simpl; autorewrite with H'_rw E0_rw ... 
    rewrite IHi; abstract lia. 
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H'_omega : forall k, H'_ omega k = S (2 * k)%nat.
Proof with auto with E0.
  intro k; rewrite H'_eq3 ...
  - replace (Canon omega (S k)) with (Fin (S k)).
    + rewrite H'_Fin; abstract lia.
    + now autorewrite with E0_rw.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H'_Plus_Fin alpha : forall i k : nat,
    H'_ (alpha + i)%e0 k = H'_ alpha (i + k)%nat.
(* Proof by induction on i *)
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H'_omega_double k : H'_ (omega * 2)%e0 k = (4 * k + 3)%nat.
Proof.
  rewrite H'_eq3; simpl Canon.
 -   ochange  (CanonS  (omega * FinS 1)%e0 k) (omega + (S k))%e0.
  + rewrite H'_Plus_Fin, H'_omega;  abstract lia.
  -  now compute.
Qed.

Lemma H'_omega_3 k : H'_ (omega * 3)%e0 k = (8 * k + 7)%nat.

Lemma H'_omega_4 k : H'_ (omega * 4)%e0 k = (16 * k + 15)%nat.

Lemma H'_omega_i i  : forall k,
    H'_ (omega * i)%e0 k = (exp2 i * k + Nat.pred (exp2 i))%nat.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H'_omega_sqr : forall k,
    H'_ (phi0  2)%e0 k = (exp2 (S k ) * (S k) - 1)%nat.
Proof.
  intro k; rewrite H'_eq3; auto with E0.
  - ochange (Canon (phi0 2) (S k)) (omega * (S k))%e0.
    +  rewrite H'_omega_i; simpl (exp2 (S k)).
       *  rewrite Nat.add_pred_r.
          -- abstract lia. 
          -- generalize (exp2_not_zero k); abstract lia.
    + cbn; f_equal; abstract lia.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H'_Omega_term_1 : alpha <> Zero -> forall k,  
    H'_ (Omega_term alpha (S i)) k =
    H'_ (Omega_term alpha i) (H'_ (phi0 alpha) k).
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H'_Omega_term (alpha : E0):
alpha <> Zero -> forall i k, 
  H'_ (Omega_term alpha i) k = iterate (H'_ (phi0 alpha)) (S i) k.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition H'_succ_fun f k := iterate f (S k) k.

Lemma H'_phi0_succ alpha  : alpha <> Zero -> forall k,
 H'_ (phi0 (Succ alpha)) k = H'_succ_fun (H'_ (phi0 alpha)) k.

Lemma H'_phi0_Si : forall i k,
 H'_ (phi0 (S i)) k = iterate H'_succ_fun i (H'_ omega) k. 
\end{Coqsrc}

\end{enumerate}
We get now a  formula for $H'_{\omega^3}$:

\begin{Coqsrc}
Lemma H'_omega_cube : forall k,
    H'_ (phi0 3)%e0 k = iterate (H'_ (phi0 2))  (S k) k.
Proof.
  intro k; rewrite <-FinS_eq, -> Fin_Succ, H'_phi0_succ; auto.
  compute; injection 1; discriminate.
Qed.
\end{Coqsrc}

@@

It seems hard to capture the complexity of this function by looking only at this
``exact'' formula. 
Let us consider a simple example: the number $H'_{\omega^3}(3)$.

\begin{Coqsrc}
Section H'_omega_cube_3.

Let f k :=   (exp2 (S k ) * (S k) - 1)%nat.

Remark R0 k :  H'_ (phi0 3)%e0 k = iterate f (S k) k.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fact F0 : H'_ (phi0 3) 3 = f (f (f (f 3))).
proof. rewrite R0; reflexivity. Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Let N := (exp2 64 * 64 - 1)%nat.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fact F1 : H'_ (phi0 3) 3 = f (f N).
Proof.
 rewrite F0; reflexivity. 
Qed.

Lemma F1_simpl : H'_ (phi0 3) 3 =
                 (exp2 (exp2 (S N) * S N) * (exp2 (S N) * S N) - 1)%nat.

\end{Coqsrc}

@@

\begin{Coqsrc}
End H'_omega_cube_3.
\end{Coqsrc}

>>>>

\input{movies/snippets/L_alpha/LFiniteOmega}

@@

\input{movies/snippets/L_alpha/LGeS}

\input{movies/snippets/L_alpha/LCorrect}

@@

Lemma \texttt{battle\_length\_std } of
Module~\href{../theories/html/hydras.Hydra.Battle_length.html}{Hydra.Battle\_length} relates the length of standard battles with the functions $L_\alpha$.

\input{movies/snippets/Battle_length/battleLengthStd}

@@

\input{movies/snippets/Hprime/HprimeDef}

\input{movies/snippets/Hprime/paraphrases}

@@

\input{movies/snippets/Hprime/HprimeFin}

@@

\input{movies/snippets/Hprime/HprimeOmega}

@@

\input{movies/snippets/Hprime/HprimePlusFin}

@@

\input{movies/snippets/Hprime/HprimeExamples}

@@

\input{movies/snippets/Hprime/HprimeOmegaSqr}

@@

\input{movies/snippets/Hprime/HprimeOmegaTerm1}

@@

\input{movies/snippets/Hprime/HprimeOmegaTerm}

@@

\input{movies/snippets/Hprime/HprimeSuccFun}
\input{movies/snippets/Hprime/HprimePhi0SI}

\item We get now a  formula for $H'_{\omega^3}$:

\input{movies/snippets/Hprime/HprimeOmegaCube}
\end{enumerate}

@@

It looks hard to capture the complexity of this function by looking only at this
``exact'' formula. 
Let us consider a simple example: the number $H'_{\omega^3}(3)$.

\input{movies/snippets/Hprime/HprimeOmegaCube3a}

@@

\input{movies/snippets/Hprime/HprimeOmegaCube3b}

@@

\input{movies/snippets/Hprime/HprimeOmegaCube3c}

@@

\input{movies/snippets/Hprime/HprimeOmegaCube3d}

@@
